---
title: "File for plotting GA and Pidiq results"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



#Step 1: make sure all the scripts are in the same directory - setwd to this directory
```{r}
rm(list=ls())


your_directory_here <- "/Users/keiranmaskell/Desktop/scripts_nightmare/1205/src"

setwd("src")

sprintf("current working directory: %s", getwd())

```


#Init script will load packages, and the fxn_repo script will load all the other functions needed
```{r}
source("init_1205.R")
source("fxn_repo 2 12_05.R")
```

#Step 2: set the parameters for the data you want to load in
#The most important parameters here are filepaths to the GA and Pidiq xlsx filepaths
#explanation of the rest of the parameters
```{r}
#paste filepath of master data file(s) WITHOUT the file extension
prms <- make.prms(flat_id = "K0074",
                        type= "single_flat",
                        GA = TRUE,
                        Pidiq = FALSE,
                        GA_master_fp = "/Users/keiranmaskell/Desktop/scripts_nightmare/1205/data/Infection_results_master",
                        GA_sheet = 1,
                        Pidiq_master_fp = "/Users/keiranmaskell/Desktop/scripts_nightmare/1205/data/Infection_results_master",
                        Pidiq_sheet = 2
                      )

```


#load in the data and calculate logcfu/cm^2 and/or PIDIQ scores
```{r}
GA_fitness <- prepare_GA_data(use_coresize_approximation = "N")
Pidiq_fitness <- prepare_Pidiq_data()

head(GA_fitness)
head(Pidiq_fitness)

```


# ```{r}

# # datalist <- get_data(prms$GA_master_fp, prms$GA_sheet, prms$Pidiq_master_fp, prms$Pidiq_sheet, GA_yn=TRUE, Pidiq_yn=TRUE)

# # datalist

# ```


#plot data from specific flats
```{r}

names(GA_fitness)

#add any treatments other than inoculum to include as variables
#extra_treatment_groups <- c("OD","Treatment..flatcol.")
extra_treatment_groups <- c("OD")

#which flats
flat_id <- c("K0074","K0075")
flat_id <- c("K0070","K0069")

#use pidiq data or not
use_pidiq <- "Y"
use_GA <- "Y"


# split_by_flatcol_yn <- "Y"
# x4_factor <- "Y"


```


#exclude certain treatments or outcomes from the data
```{r}

#treatments_to_exclude <- list(Inoculum="plusminus_PtoDC3000::EV_and_D36E::EV",insufficient_material="Y")
treatments_to_exclude <- list(Inoculum="plusminus_PtoDC3000::EV_and_D36E::EV")


if(use_GA=="Y"){
data_subset <- GA_fitness
for(i in 1:length(treatments_to_exclude)){
  data_subset <- data_subset[data_subset[,names(treatments_to_exclude[i])]!=treatments_to_exclude[[i]],]
  
  }
}

if(use_pidiq=="Y"){
pidiq_subset <- Pidiq_fitness
for(i in 1:length(treatments_to_exclude)){
  pidiq_subset <- pidiq_subset[pidiq_subset[,names(treatments_to_exclude[i])]!=treatments_to_exclude[[i]],]
  
  }
}


``` 


#filter the data by the chosen flats, and split into data with 4 cores and data without 4 cores, as pooling cores can affect the overall outcome
#aggregate the technical reps to get average for each biological replicate
#merge pidiq and GA data if both are being used, and remove data with empties for either pidiq or GA
#If plotting GA only or Pidiq only, you can keep the data with empties for the other
```{r}

# #check for duplicates 
# duplicates <- pidiq_subset[duplicated(pidiq_subset[, c("Flat","Inoculum", "Plant")]),]
# print(duplicates)


if(nrow(data_subset[data_subset$Flat %in% flat_id,])==0){
  use_GA <- "N"
}

if(nrow(pidiq_subset[pidiq_subset$Flat %in% flat_id,])==0){
  use_pidiq <- "N"
}


if(use_GA=="Y"){

  #only the targeted flats
  data_subset <- data_subset[data_subset$Flat %in% flat_id,]

  #split into full sample and suspects (less than 4 cores or insufficient material to get proper cores) 
  #-> the latter group will be excluded from statistical analysis
  data_subset_fullsample <- data_subset[data_subset$insufficient_material=="N"|as.numeric(data_subset$core.number)==4,][order(data_subset$Inoculum[data_subset$insufficient_material=="N"|as.numeric(data_subset$core.number)==4]),]
  data_subset_suspects <- data_subset[data_subset$insufficient_material=="Y"|as.numeric(data_subset$core.number)!=4,][order(data_subset$Inoculum[data_subset$insufficient_material=="Y"|as.numeric(data_subset$core.number)!=4]),]

  #aggregate the technical reps to get average for each biological replicate
  data_subset_fullsample_agg <- aggregate(data_subset_fullsample[,"log.CFU.cm2"], by = data_subset_fullsample[,c("Flat","Inoculum","Plant",extra_treatment_groups)], FUN = mean)
  data_subset_suspects_agg <- aggregate(data_subset_suspects[,"log.CFU.cm2"], by = data_subset_suspects[,c("Flat","Inoculum","Plant",extra_treatment_groups)], FUN = mean)
  
  #both GA and Pidiq
  if(use_pidiq=="Y"){
    pidiq_subset <- pidiq_subset[pidiq_subset$Flat %in% flat_id,]

    merged_data_fullsample <- merge(data_subset_fullsample_agg, pidiq_subset[, c("Flat","Inoculum", "Plant", "Arcsine.transformed.data")], 
                     by = c("Flat","Inoculum", "Plant"), all.x = TRUE)

    merged_data_suspects <- merge(data_subset_suspects_agg, pidiq_subset[, c("Flat","Inoculum", "Plant", "Arcsine.transformed.data")],
                      by = c("Flat","Inoculum", "Plant"), all.x = TRUE)

  #only GA
  }else{
    merged_data_fullsample <- data_subset_fullsample_agg
    merged_data_suspects <- data_subset_suspects_agg
  
  }
}else{
  #only Pidiq
  if(use_pidiq=="Y"){
    pidiq_subset <- pidiq_subset[pidiq_subset$Flat %in% flat_id,]
    merged_data_fullsample <- pidiq_subset
    merged_data_suspects <- pidiq_subset
  }else{
    stop("No data to plot")
  
  }
}

#remove data with empties for either pidiq or GA - if you want to keep them plot GA only or Pidiq only
# Find rows in pidiq_subset that don't have a match in data_subset - these are the biological replicates 
no_match <- merged_data_fullsample[is.na(merged_data_fullsample$Arcsine.transformed.data)|is.na(merged_data_fullsample$x),]
print(no_match)
if(nrow(no_match)>0){
  filtered_df_fullsample <- merged_data_fullsample[!is.na(merged_data_fullsample$Arcsine.transformed.data) & !is.na(merged_data_fullsample$x), ]
}else{
  filtered_df_fullsample <- merged_data_fullsample
}

no_match <- merged_data_suspects[is.na(merged_data_suspects$Arcsine.transformed.data)|is.na(merged_data_suspects$x),]
print(no_match)
if(nrow(no_match)>0){
  filered_df_suspects <- merged_data_suspects[!is.na(merged_data_suspects$Arcsine.transformed.data) & !is.na(merged_data_suspects$x), ]
}else{
  filtered_df_suspects <- merged_data_suspects
}


```


# the_dpi <- "4"

# print("date infected")
# print(unique(as.Date(data_subset$Date.infected[data_subset$Flat==flat_id])))

# print("date harvested")
# harvest_date <- unique(paste0(
#    substring(unique(as.Date(data_subset$Date.infected[data_subset$Flat==flat_id])),0,8),
#    as.character(as.numeric(data_subset$Days.infected.at.time.of.harvest) + as.numeric(substring(unique(as.Date(data_subset$Date.infected[data_subset$Flat==flat_id])),9,nchar(as.character((unique(as.Date(data_subset$Date.infected[data_subset$Flat==flat_id]))))))))
# ))
# print(harvest_date)

#remove the extra columns from the df
```{r}

if(use_GA=="Y"){
  if(use_pidiq=="Y"){

    filtered_df_fullsample <- filtered_df_fullsample %>%
    dplyr::select(Inoculum,extra_treatment_groups,x,Arcsine.transformed.data)

    filtered_df_suspects <- filtered_df_suspects %>%
    dplyr::select(Inoculum,extra_treatment_groups,x,Arcsine.transformed.data)

  }else{

    filtered_df_fullsample <- filtered_df_fullsample %>%
    dplyr::select(Inoculum,extra_treatment_groups,x)

    filtered_df_suspects <- filtered_df_suspects %>%
    dplyr::select(Inoculum,extra_treatment_groups,x)

    }
  }else{
    if(use_pidiq=="Y"){
      filtered_df_fullsample <- filtered_df_fullsample %>%
      dplyr::select(Inoculum,extra_treatment_groups,Arcsine.transformed.data)

      filtered_df_suspects <- filtered_df_suspects %>%
      dplyr::select(Inoculum,extra_treatment_groups,Arcsine.transformed.data)
    }else{
    stop("No data to plot")
  }}

```


#significance groups
```{r}

for(i in 1:length(extra_treatment_groups)){
  filtered_df_fullsample[,extra_treatment_groups[i]] <- as.factor(filtered_df_fullsample[,extra_treatment_groups[i]])
  filtered_df_suspects[,extra_treatment_groups[i]] <- as.factor(filtered_df_suspects[,extra_treatment_groups[i]])
  }

#data_subset_fullsample_agg$OD <- as.numeric(data_subset_fullsample_agg$OD)


#ensure that the specified extra_treatment_groups actually have different levels
extra_treatment_groups <- extra_treatment_groups[length(unique(filtered_df_fullsample[,extra_treatment_groups]))>1]

if(length(extra_treatment_groups)>0){
  interaction_terms <- paste("Inoculum", extra_treatment_groups, sep = "*", collapse = "+")
  }else{
    print("removing extra treatment groups, as there is only one level for all specified")
    interaction_terms <- "Inoculum"
  }
formula <- as.formula(paste("x ~", interaction_terms))
model <- aov(formula, data = filtered_df_fullsample)
summary(model)

tukey_result <- TukeyHSD(model)
cld <- multcompLetters4(model,tukey_result)
cld


if(length(extra_treatment_groups)>0){
  interaction_group <- paste('Inoculum', extra_treatment_groups, sep = ':')
  Letters <- cld[[interaction_group]]$Letters
  print(Letters)
  }else{
    print("No extra treatment groups to compare")
    Letters <- cld$Inoculum$Letters
  }
  print(Letters)

```
# interaction_group <- paste('Inoculum', extra_treatment_groups, sep = ':')
# Letters <- cld[[interaction_group]]$Letters
# print(Letters)




```{r}

filtered_df_fullsample$box_colour <- ifelse(data_subset_fullsample_agg$Inoculum=="PtoDC3000::EV","gold",ifelse(data_subset_fullsample_agg$Inoculum=="D36E::EV","darkgreen","grey"))
filtered_df_suspects$box_colour <- ifelse(data_subset_suspects_agg$Inoculum=="PtoDC3000::EV","gold",ifelse(data_subset_suspects_agg$Inoculum=="D36E::EV","darkgreen","grey"))

filtered_df_fullsample$point_colour <- 'black'
filtered_df_suspects$point_colour <- 'red'

treatment_order <- c("D36E::EV", 
                     unique(as.character(filtered_df_fullsample$Inoculum)[filtered_df_fullsample$Inoculum!="PtoDC3000::EV"&filtered_df_fullsample$Inoculum!="D36E::EV"]), 
                     "PtoDC3000::EV")

factor_Letters <- factor(names(Letters), levels = desired_order)
ordered_Letters <- Letters[order(factor_Letters)]



filtered_df_fullsample <- filtered_df_fullsample[order(factor(filtered_df_fullsample$Inoculum, levels = treatment_order)),]
filtered_df_suspects <- filtered_df_suspects[order(factor(filtered_df_suspects$Inoculum, levels = treatment_order)),]



anova_data <- summary(model)
# # Calculate p-value from the ANOVA result
p_value <- anova_data[[1]]$Pr[1]

if(round(p_value,1)=="0"){
   for(i in 1:20){
      if(round(p_value,i)!=0){
         round_lvl <- i
         break()
      }
   }
}else{
   round_lvl <- 1
}


# print("date infected")
# print(unique(as.Date(data_subset$Date.infected[data_subset$Flat==flat_id])))

# print("date harvested")
# harvest_date <- unique(paste0(
#    substring(unique(as.Date(data_subset$Date.infected[data_subset$Flat==flat_id])),0,8),
#    as.character(as.numeric(data_subset$Days.infected.at.time.of.harvest) + as.numeric(substring(unique(as.Date(data_subset$Date.infected[data_subset$Flat==flat_id])),9,nchar(as.character((unique(as.Date(data_subset$Date.infected[data_subset$Flat==flat_id]))))))))
# ))
# print(harvest_date)




```




#plot the GA data
```{r}

#ordered_Letters

the_dpi <- ifelse(length(unique(data_subset$Days.infected.at.time.of.harvest))==1,unique(data_subset$Days.infected.at.time.of.harvest),paste(unique(data_subset$Days.infected.at.time.of.harvest),collapse=", "))

color_mapping <- c("darkgreen",rep("grey",length(unique(filtered_df_fullsample$Inoculum))-2),"gold")

filtered_df_fullsample$Inoculum <- factor(filtered_df_fullsample$Inoculum, levels = treatment_order)


plot <- ggplot(filtered_df_fullsample, aes(x=Inoculum, y=x)) +
  geom_boxplot(aes(fill=Inoculum),outlier.shape = NA) +
  scale_fill_manual(values = color_mapping) +
  geom_point(aes(color=Inoculum), position=position_jitterdodge(), size=3) +
  scale_color_manual(values = rep("black",nrow(filtered_df_fullsample))) +
#   geom_text(data = data.frame(x = unique(filtered_df_fullsample$Inoculum), y = a$stats[nrow(a$stats),], Letters = ordered_Letters),
#             aes(x = x, y = y+0.3, label = Letters), color = "black") +
geom_text(data = data.frame(x = levels(filtered_df_fullsample$Inoculum), Letters = ordered_Letters),
            aes(x = x, y = (max(filtered_df_fullsample$x)+0.4), label = Letters), color = "black", size=8) +

  xlab("Treatment") +
  ylab("log(CFU/cm^2)") +
  ylim(0, max(filtered_df_fullsample$x) + 1.0) +
  labs(
    title = sprintf("Flat %s Bacterial fitness results at %s dpi", paste(flat_id,collapse=", "), the_dpi),
    subtitle = paste("p =", round(p_value, round_lvl))
  ) + 
  theme(
    axis.title.x = element_text(size = 20),
    axis.title.y = element_text(size = 20),
    plot.title = element_text(size = 20),
    plot.subtitle = element_text(size = 20),
    axis.text = element_text(size = 20),
    legend.position = "none"
  )

plot <- plot +
  geom_point(data = filtered_df_suspects, color="red", position = position_jitter(), alpha = 1, size = 3.0)
plot

```

#save the plot
```{r}
Plot_fp <- "your filepath to save the plot here"

ggsave(Plot_fp, plot, width = 10, height = 10, units = "in", dpi = 300)

```

#plot the Pidiq data
```{r}

#First recalculate the significance groups using the combined data (no need to remove plants with insufficient material, as this is not relevant for Pidiq)
#we'll still plot those plants as red data points
the_dpi <- ifelse(length(unique(data_subset$Days.infected.at.time.of.harvest))==1,unique(data_subset$Days.infected.at.time.of.harvest),paste(unique(data_subset$Days.infected.at.time.of.harvest),collapse=", "))

color_mapping <- c("darkgreen",rep("grey",length(unique(filtered_df_fullsample$Inoculum))-2),"gold")

combined_sets <- rbind(filtered_df_fullsample,filtered_df_suspects)
combined_sets$Inoculum <- factor(combined_sets$Inoculum, levels = treatment_order)


formula <- as.formula(paste("Arcsine.transformed.data ~", interaction_terms))
model <- aov(formula, data = combined_sets)
summary(model)

tukey_result <- TukeyHSD(model)
cld <- multcompLetters4(model,tukey_result)
cld


if(length(extra_treatment_groups)>0){
  interaction_group <- paste('Inoculum', extra_treatment_groups, sep = ':')
  Letters <- cld[[interaction_group]]$Letters
  print(Letters)
  }else{
    print("No extra treatment groups to compare")
    Letters <- cld$Inoculum$Letters
  }
  print(Letters)


treatment_order <- c("D36E::EV", unique(filtered_df_fullsample$Inoculum[filtered_df_fullsample$Inoculum!="PtoDC3000::EV"&filtered_df_fullsample$Inoculum!="D36E::EV"]), "PtoDC3000::EV")
factor_Letters <- factor(names(Letters), levels = desired_order)
ordered_Letters <- Letters[order(factor_Letters)]

anova_data <- summary(model)
# # Calculate p-value from the ANOVA result
p_value <- anova_data[[1]]$Pr[1]

if(round(p_value,1)=="0"){
   for(i in 1:20){
      if(round(p_value,i)!=0){
         round_lvl <- i
         break()
      }
   }
}else{
   round_lvl <- 1
}




plot <- ggplot(combined_sets, aes(x=Inoculum, y=Arcsine.transformed.data)) +
  geom_boxplot(aes(fill=Inoculum),outlier.shape = NA) +
#   geom_point(data=filtered_df_fullsample, position=position_jitter(), size=3) +
#     scale_color_manual(values = rep("black",nrow(filtered_df_fullsample)))+
  scale_fill_manual(values = color_mapping) +
#   geom_text(data = data.frame(x = unique(combined_sets$Inoculum), y = a$stats[nrow(a$stats),], Letters = ordered_Letters),
#             aes(x = x, y = y+0.3, label = Letters), color = "black") +
geom_text(data = data.frame(x = levels(combined_sets$Inoculum), Letters = ordered_Letters),
            aes(x = x, y = (max(combined_sets$Arcsine.transformed.data)+0.1), label = Letters), color = "black", size=8) +

  xlab("Treatment") +
  ylab("Arcsin(% yellow)") +
  ylim(0, max(combined_sets$Arcsine.transformed.data) + 0.2) +
  labs(
    title = sprintf("Flat %s Bacterial fitness results at %s dpi", paste(flat_id,collapse=", "), the_dpi),
    subtitle = paste("p =", round(p_value, round_lvl))
  ) + 
  theme(
    axis.title.x = element_text(size = 20),
    axis.title.y = element_text(size = 20),
    plot.title = element_text(size = 20),
    plot.subtitle = element_text(size = 20),
    axis.text = element_text(size = 20),
    legend.position = "none"
  )
  plot
plot <- plot +
    geom_point(data=filtered_df_fullsample, position=position_jitter(), size=3) +
    scale_color_manual(values = rep("black",nrow(filtered_df_fullsample))) +
    geom_point(data = filtered_df_suspects, color="red", position = position_jitter(), alpha = 1, size = 3.0)

plot  

```


#save the Pidiq plot
```{r}

Plot_fp <- "your filepath to save the plot here"
ggsave(Plot_fp, plot, width = 10, height = 10, units = "in", dpi = 300)

```

#plot Pidiq by GA
```{r}
install.packages("ellipse")
library(ellipse)
color_sequence <- colorRampPalette(c("darkgreen", "gold"))(100)

#color_mapping <- c("darkgreen",rep("grey",length(unique(filtered_df_fullsample$Inoculum))-2),"gold")
#pidiq_colour <- color_sequence[as.numeric(100*round(filtered_df_fullsample$Arcsine.transformed.data,2))]

##filtered_df_fullsample$pidiq_colour <- color_sequence[as.numeric(cut(filtered_df_fullsample$Arcsine.transformed.data, breaks = 100))]
filtered_df_fullsample$pidiq_colour <- color_sequence[as.numeric(100*round(filtered_df_fullsample$Arcsine.transformed.data,2))]
filtered_df_suspects$pidiq_colour <- color_sequence[as.numeric(100*round(filtered_df_suspects$Arcsine.transformed.data,2))]
combined_sets$pidiq_colour <- color_sequence[as.numeric(100*round(combined_sets$Arcsine.transformed.data,2))]




mean_pidiqs <- combined_sets %>%
    group_by(Inoculum) %>%
    summarise(mean_pidiq = 100*round(mean(Arcsine.transformed.data),2))

mean_pidiqs_colours <- color_sequence[as.numeric(mean_pidiqs$mean_pidiq)]

combined_sets$color <- mean_pidiqs_colours[as.numeric(as.factor(combined_sets$Inoculum))]


plot <- ggplot(combined_sets, aes(x=Arcsine.transformed.data, y=x)) +
    geom_point(data=filtered_df_suspects, size=8, color="red", shape=19)+

  xlab("Asin(% yellow)") +
  ylab("log(CFU/cm^2)") +
  ylim(0, max(combined_sets$x) + 0.5) +
  labs(
    title = sprintf("Flat %s Bacterial fitness results at %s dpi", paste(flat_id,collapse=", "), the_dpi),
    subtitle = paste("p =", round(p_value, round_lvl))
  ) + 
  theme(
    axis.title.x = element_text(size = 20),
    axis.title.y = element_text(size = 20),
    plot.title = element_text(size = 20),
    plot.subtitle = element_text(size = 20),
    axis.text = element_text(size = 20),
    legend.text = element_text(size = 20)

  )

plot

plot<- plot+  geom_point(aes(shape=Inoculum), size=5, color=pidiq_colour) +
  #stat_ellipse(aes(group=Inoculum), level=0.95, geom="path") +
stat_ellipse(aes(group=Inoculum, fill=Inoculum, color=Inoculum), level=0.95, geom="polygon", alpha=0.1) +
scale_fill_manual(values = mean_pidiqs_colours) +
scale_color_manual(values = rep("grey",length(unique(combined_sets$Inoculum))))

plot


```


#save the Pidiq x GA plot
```{r}

Plot_fp <- "your filepath to save the plot here"
ggsave(Plot_fp, plot, width = 10, height = 10, units = "in", dpi = 300)

```